---
title: "DATA 606 - Lab 7 - Inference for Numerical Data"
author: "Preston Peck"
date: "11/1/2021"
output:
  pdf_document: default
  html_document: default
---

# Inference for Numerical Data
<https://htmlpreview.github.io/?https://github.com/jbryer/DATA606/blob/master/inst/labs/Lab7/Lab7_inf_for_numerical_data.html>

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
library(infer)
library(psych)
library(Rmisc)
library(DescTools)
```

```{r}
set.seed(37)

data('yrbss', package='openintro')
yrbss
```

### Exercise 1
There are 13583 cases (rows)
```{r}
yrbss %>%
  glimpse

yrbss %>%
  nrow
```

### Exercise 2
There are valid 12579 entries and 1004 NA's
```{r}
yrbss$weight %>%
  describe

yrbss$weight %>%
  summary
```

### Exercise 3
Those working out more than 2 days a week weight more, and that might be accredited to increased muscle mass
```{r}
yesLabel <- "yes"
noLabel <- "no"



yrbss <- yrbss %>% 
  mutate(physical_3plus = ifelse(yrbss$physically_active_7d > 2, yesLabel, noLabel))

boxplot(yrbss$weight ~ yrbss$physical_3plus, ylab = "Weight", xlab = "Physical 3+ Days")


no_physical_3plus <- (yrbss %>%
  filter(physical_3plus == noLabel))

no_physical_3plus$weight %>%
  summary

no_physical_3plus$weight %>%
  describe


yes_physical_3plus <- (yrbss %>%
  filter(physical_3plus == yesLabel))

yes_physical_3plus$weight %>%
  summary

yes_physical_3plus$weight %>%
  describe


yrbss %>%
  group_by(physical_3plus) %>%
  dplyr::summarise(mean_weight = mean(weight, na.rm = TRUE))
```

### Exercise 4
There are 3 conditions for interference: Random, Normal, Independence \
The sample is random OK \
The sample skew is normal ? \
-> If the skew is not normal (right skew) or it has an unknown parent population, the sample size must be large (n >= 30) OK \
The sample is independent (n <= 10% of population — 13583 / 4.9 million high school students from 1991-2019) OK

```{r}
yrbss %>%
  group_by(physical_3plus) %>%
  filter(!is.na(weight)) %>%
  dplyr::count(physical_3plus)


ggplot(data = yrbss, aes(x = weight)) +
    geom_histogram()

ggplot(data = no_physical_3plus, aes(x = weight)) +
    geom_histogram()

ggplot(data = yes_physical_3plus, aes(x = weight)) +
    geom_histogram()
```

### Exercise 5
H0: There is no correlation between weight and number of days worked out in a week \
HA: There is a correlation between weight and number of days worked out in a week

### Exercise 6
No values are greater than obs_stat \
Since the p-value is less than 0.05, we can reject the null hypothesis
```{r}
obs_diff <- yrbss %>%
  specify(weight ~ physical_3plus) %>%
  calculate(stat = "diff in means", order = c("yes", "no"))

obs_diff

null_dist <- yrbss %>%
  specify(weight ~ physical_3plus) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in means", order = c("yes", "no"))

null_dist

ggplot(data = null_dist, aes(x = stat)) +
  geom_histogram()

null_dist %>%
  get_p_value(obs_stat = obs_diff, direction = "two_sided")
```

### Exercise 7
We're 95% sure the difference of averages of all weights of those who don't work 3 days and greater and those that do falls somewhere between -2.424441 and -1.124728
```{r}
DATA606::inference(y = yrbss$weight, 
                   x = yrbss$physical_3plus, 
                   null = 0,
                   est = "mean",
                   type = "ht",
                   alternative = "twosided", 
                   method = "theoretical")

oilabs::inference(y = weight, 
                  x = physical_3plus,
                  data = yrbss,
                  null = 0,
                  statistic = "mean",
                  type = "ht",
                  alternative = "twosided", 
                  method = "theoretical")

t.test(data = yrbss, weight ~ physical_3plus)

yes_physical_3plus$weight %>%
  t.test

no_physical_3plus$weight %>%
  t.test
```

### Exercise 8
We're 95% sure the average of all heights of those who work 3 days and greater and those that don't falls somewhere between 1.689411 and 1.693071
```{r}
yrbss$height %>%
  summary

yrbss_heights <- yrbss %>%
  filter(!is.na(height))



physical_3plus_grouped <- yrbss_heights %>%
  group_by(physical_3plus)
  
physical_3plus_grouped


physical_3plus_grouped_means <- physical_3plus_grouped %>%
  dplyr::summarise(mean(height))
  
physical_3plus_grouped_means

yrbss_mean <- mean(yrbss_heights$height)

yrbss_mean


physical_3plus_grouped_counts <- physical_3plus_grouped %>%
  dplyr::count(physical_3plus)

physical_3plus_grouped_counts

yrbss_count <- nrow(yrbss_heights)

yrbss_count


physical_3plus_grouped_sds <- physical_3plus_grouped %>%
  dplyr::summarise(sd(height))

physical_3plus_grouped_sds

yrbss_sd <- sd(yrbss_heights$height)

yrbss_sd


z = 1.96


upperCI <- yrbss_mean + z * (yrbss_sd / sqrt(yrbss_count))

upperCI

lowerCI <- yrbss_mean - z * (yrbss_sd / sqrt(yrbss_count))

lowerCI



yrbss_heights$height %>%
  summary


yrbss_heights$height %>%
  t.test

ciBounds <- yrbss_heights$height %>%
  CI

ciBounds

meanCIBounds <- yrbss_heights$height %>%
  MeanCI

meanCIBounds



c(upperCI, ciBounds[[1]], meanCIBounds[[3]])

c(lowerCI, ciBounds[[3]], meanCIBounds[[2]])

paste("Interval:", ciBounds[[1]] - ciBounds[[3]])



DATA606::inference(y = yrbss_heights$height, 
                   type = "ci", 
                   est = "mean",
                   alternative = "twosided", 
                   method = "theoretical")

oilabs::inference(y = height, 
                 data = yrbss_heights,
                 null = 0,
                 statistic = "mean",
                 type = "ci",
                 alternative = "twosided", 
                 method = "theoretical")
```

### Exercise 9
The range is smaller at 90% confidence because we're less confident, which means by reducing the range, the true average is less likely to fall within it
```{r}
z <- 1.645



yrbss_heights$height %>%
  t.test(conf.level = 0.90)

ciBounds <- yrbss_heights$height %>%
  CI(ci = 0.90)

ciBounds


meanCIBounds <- yrbss_heights$height %>%
  MeanCI(conf.level = 0.90)

meanCIBounds



paste("Upper bound:", ciBounds[[1]])

paste("Lower bound:", ciBounds[[3]])

paste("Interval:", ciBounds[[1]] - ciBounds[[3]])



DATA606::inference(y = yrbss_heights$height, 
                   type = "ci", 
                   est = "mean", 
                   alternative = "twosided", 
                   method = "theoretical", 
                   conflevel = 0.90)

oilabs::inference(y = height, 
                  data = yrbss_heights,
                  null = 0,
                  statistic = "mean",
                  type = "ci",
                  alternative = "twosided", 
                  method = "theoretical")
```

### Exercise 10
H0: There is no correlation between height and number of days worked out in a week \
HA: There is a correlation between height and number of days worked out in a week \
Since the p-value is less than 0.05, we can reject the null hypothesis
```{r}
yes_physical_3plus_yrbss_heights <- (yrbss_heights %>%
  filter(physical_3plus == yesLabel))

yes_physical_3plus_yrbss_heights$height %>%
  summary

yes_physical_3plus_yrbss_heights$height %>%
  MeanCI



no_physical_3plus_yrbss_heights <- (yrbss_heights %>%
  filter(physical_3plus == noLabel))

no_physical_3plus_yrbss_heights$height %>%
  summary

no_physical_3plus_yrbss_heights$height %>%
  MeanCI



DATA606::inference(y = yrbss_heights$height, 
                   x = yrbss_heights$physical_3plus, 
                   null = 0,
                   est = "mean",
                   order = c(yesLabel, noLabel),
                   type = "ht",
                   alternative = "twosided", 
                   method = "theoretical")

oilabs::inference(y = height, 
                  x = physical_3plus, 
                  data = yrbss_heights,
                  null = 0,
                  statistic = "mean",
                  order = c(yesLabel, noLabel),
                  type = "ht",
                  alternative = "twosided", 
                  method = "theoretical")
```

### Exercise 11
There are 8 options for hours_tv_per_school_day (one of which is NA)
```{r}
yrbss %>%
  group_by(hours_tv_per_school_day) %>%
  dplyr::summarise(n())
```

### Exercise 12
H0: There is no correlation between weight and hours of sleep on a school night \
HA: There is a correlation between weight and hours of sleep on a school night \
Since the p-value is less than 0.05, we can reject the null hypothesis
```{r}
yrbss_weight_school_night_hours_sleep <- yrbss %>%
  filter(!is.na(weight)) %>%
  filter(!is.na(school_night_hours_sleep))



oilabs::inference(y = weight, 
                  x = school_night_hours_sleep, 
                  data = yrbss_weight_school_night_hours_sleep,
                  null = 0,
                  statistic = "mean",
                  order = c(yesLabel, noLabel),
                  type = "ht",
                  alternative = "greater", 
                  method = "theoretical")
```